% references
\documentclass[../main.tex]{subfiles}

\begin{document}

% Chicago style

\begin{thebibliography}{9}
\bibitem{Beardon}
Beardon, Alan F. 2000. \textit{Iteration Of Rational Functions}. New York, NY: Springer.

\bibitem{Brjuno}
Brjuno, Alexander D. 1965. \textit{Convergence of transformations of differential equations to normal forms}. Dokl. Akad. Nauk USSR 165, 987-989
(Soviet Math. Dokl., 1536-1538).

\bibitem{Buff}
Buff, Xavier and Ch\'eritat, Arnaud. 2004. \textit{Upper Bound For The Size Of Quadratic Siegel Disks}. Inventiones Mathematicae 156 (1): 1-24. doi:10.1007/s00222-003-0331-6.

\bibitem{Gamelin}
Carleson, Lennart and Gamelin, Theodore W. 2013. \textit{Complex Dynamics}. New York, NY: Springer.

\bibitem{Cremer}
Cremer, Hubert. 1938. \textit{Über die Häufigkeit der Nichtzentren}, Math. Ann. 115, no. 1, 573--580. MR MR15132

\bibitem{ecalle}
Écalle, Jean. 1981. \textit{Les fonctions resurgentes, volumes I–III}. Publ. Math. Orsay, France.

\bibitem{Fernandez}
Fernandez, Jaume A. 2012. \textit{Siegel’s Theorem: An Application Of KAM Theory To Complex Dynamics}. Presentation, Uppsala University, 2012.

\bibitem{Fremlin}
Fremlin, David. 2011. \textit{Measure Theory.} Avaliable from: \texttt{https://www1.essex.ac.uk/maths/people/fremlin/mt.htm}. Accessed June 2020.

\bibitem{Geyer}
Geyer, Lukas. 2016. \textit{Topics In Mathematics Complex Dynamics}. Lecture Notes, 2016.

\bibitem{shenko}
Iliashenko, IU. S, and S Yakovenko. 2008. \textit{Lectures On Analytic Differential Equations}. Providence, Rhode Island: American Mathematical Society.

\bibitem{Knopp}
Knopp, Konrad. 1990. \textit{Theory and Application of Infinite Series.} Dover Publications.

\bibitem{McMullen}
McMullen, Curtis T. 1994. \textit{Complex Dynamics and Renormalization.} Available from: \texttt{http://people.math.harvard.edu/~ctm/papers/home/text/papers/real/book.pdf}. Accessed June 2020.

\bibitem{Smale}
Miles-Leighton, Hayley, and Pilgrim, Kevin M. 2012. \textit{Smale’s Mean Value Conjecture And Complex Dynamics}. Computational Methods And Function Theory 12 (2): 559-563. doi:10.1007/bf03321844.

\bibitem{Milnornotes}
Milnor, John W. 2005. \textit{Dynamics In One Complex Variable}. Presentation, Institute for Mathematical Sciences, SUNY, Stony Brook NY.

\bibitem{milnor}
Milnor, John W. 2006. \textit{Dynamics In One Complex Variable}. Princeton, N.J.: Princeton University Press.

\bibitem{Perez-Marco}
Perez-Marco, Ricardo. 1990. \textit{Sur la dynamique des germes de diffeomorphismes holomorphes de $(\C,O)$ et des diffcomorphismes analytiques du cercle}
These, Paris-Sud.

\bibitem{Rees}
Rees, M. 1942. \textit{Ergodic rational maps with dense critical point forward orbit.} Ergod. Th. \& Dynam. Sys.4, 311–322.



\bibitem{Siegel}
Siegel, Carl L. 1942. \textit{Iteration of analytic functions}, Ann. of Math. (2) 43, 607–612. MR 7044



\bibitem{stoll}
Stoll, Danny. 2020. \textit{A Brief Introduction To Complex Dynamics}. University of Chicago. Accessed June 8.

\bibitem{Yoccoz}
Yoccoz, Jean-Christophe. 1988. \textit{Linearieation des germes de diffeomorphismes holomorphes de $(\C, 0)$}, C. R. Acad. Sci. Paris 306, 55-58.

\end{thebibliography}

\end{document}